Local Spectral Clustering of Density Upper Level Sets

### 2019-11-03 arxiv/local_spectral_density_clustering_arxiv.tex
few typos fixed

To facilitate comparisons between our upper and lower bounds, assume $\frac{1}{4}\sigma \leq \frac{1}{40}\rho$ and set $r = \frac{1}{4}\sigma$. 
For some universal constant $c > 0$, the following statements each hold with high probability:
\begin{itemize}
	\item If the user-specified parameters satisfy~\eqref{eqn: initialization}, and additionally for some $b > 0$
	\begin{equation*}
	\epsilon < c\frac{b\sigma^2}{\rho^2 \log^2\left(\frac{\Lambda_{\sigma}}{\lambda_{\sigma}^2\sigma}\right)}
	\end{equation*}
	then $\Delta(\widehat{C}, \mathcal{C}^{(1)}[\Xbf]) \leq b \cdot  \vol_{n,r}(\mathcal{C}^{(1)}[\Xbf])$.
	\item If the user-specified parameters are given as in Theorem~\ref{thm:ppr_lb}, and 
	\begin{equation*}
	\epsilon > c\left({\frac{\sigma}{\rho}} \log\left(\frac{\rho \sigma}{\epsilon^2 r^2}\right)\right)^{1/4}
	\end{equation*}
	then $\Delta(\widehat{C}, \mathcal{C}^{(1)}[\Xbf]) \geq \frac{1}{8} \vol_{n,r}(\mathcal{C}^{(1)}[\Xbf])$.
\end{itemize}

@@

\paragraph{Empirical behavior of PPR.} In Figure \ref{fig:fig2}, to drive home
the main implications of Theorems \ref{thm:volume_ssd} and
\ref{thm: consistent_recovery_of_density_clusters}, we show the
behavior of PPR, normalized cut, and the density clustering algorithm of
\citet{chaudhuri2010} on the well known ``two moons'' dataset (with added 2d 
Gaussian noise), considered a prototypical success story for spectral clustering
algorithms. The first column shows the empirical density clusters $\Cset[\Xbf]$
and $\Cset'[\Xbf]$ for a particular threshold $\lambda$ of the density function; the 
second column shows the cluster recovered by PPR; the third column shows the
global minimum normalized cut, computed according to the algorithm of
\citet{szlam2010}; and the last column shows a cut of the density cluster tree
estimator of \citet{chaudhuri2010}.  We can see the degrading ability of PPR to
recover density clusters as the two moons become less well-separated. Of 
particular interest is the fact that PPR fails to recover one of the moons even
when normalized cut still succeeds in doing so. Additionally, we note that the Chaudhuri-Dasgupta algorithm succeeds even when both PPR and normalized cut fail.  While our main message was
that PPR recovers geometrically well-conditioned density clusters, it would be  
interesting to establish that it \emph{only} recovers such clusters, a direction
for future work.

>>>>

To facilitate comparisons between our upper and lower bounds, assume $\frac{1}{4}\sigma \leq \frac{1}{40}\rho$ and set $r = \frac{1}{4}\sigma$. 
For some universal constant $c > 0$, the following statements each hold with high probability:
\begin{itemize}
	\item If the user-specified parameters satisfy~\eqref{eqn: initialization}, and additionally for some $b > 0$
	\begin{equation*}
	\epsilon < c\frac{b\sigma^2}{\rho^2 \log^2\left(\frac{\Lambda_{\sigma}}{\lambda_{\sigma}^2\sigma}\right)}
	\end{equation*}
	then $\Delta(\widehat{C}, \Csig^{(1)}[\Xbf]) \leq b \cdot  \vol_{n,r}(\Csig^{(1)}[\Xbf])$.
	\item If the user-specified parameters are given as in Theorem~\ref{thm:ppr_lb}, and 
	\begin{equation*}
	\epsilon > c\left({\frac{\sigma}{\rho}} \log\left(\frac{\rho \sigma}{\epsilon^2 r^2}\right)\right)^{1/4}
	\end{equation*}
	then $\Delta(\widehat{C}, \Csig^{(1)}[\Xbf]) \geq \frac{1}{8} \vol_{n,r}(\Csig^{(1)}[\Xbf])$.
\end{itemize}

@@

\paragraph{Empirical behavior of PPR.} In Figure \ref{fig:fig2}, to drive home
the main implications of Theorems \ref{thm:volume_ssd} and
\ref{thm: consistent_recovery_of_density_clusters}, we show the
behavior of PPR, normalized cut, and the density clustering algorithm of
\citet{chaudhuri2010} on the well known ``two moons'' dataset (with added 2d 
Gaussian noise), considered a prototypical success story for spectral clustering
algorithms. The first column shows the empirical density clusters $\Cset[\Xbf]$
and $\Cset'[\Xbf]$ for a particular threshold $\lambda$ of the density function; the 
second column shows the cluster recovered by PPR; the third column shows the
global minimum normalized cut, computed according to the algorithm of
\citet{szlam2010}; and the last column shows a cut of the density cluster tree
estimator of \citet{chaudhuri2010}.  We can see the degrading ability of PPR to
recover density clusters as the two moons become less well-separated. Of 
particular interest is the fact that PPR fails to recover one of the moons even
when normalized cut still succeeds in doing so. Additionally, we note that the Chaudhuri-Dasgupta algorithm succeeds even when both PPR and normalized cut fail.  This supports our main message, which is that PPR recovers only geometrically well-conditioned density clusters.
